An exposition on the propriety of restricted Boltzmann machines

### 2016-06-22 writing/cut_text.Rmd
beef up instability

>>>>

```{r instab_plots, fig.cap="\\label{fig:instab_plots}A numerical experiment looking at $(\\max\\limits_{\\boldsymbol x_N \\in \\mathcal{X}_N}[Q(\\boldsymbol x_N)] - \\min\\limits_{\\boldsymbol x_N \\in \\mathcal{X}_N}[Q(\\boldsymbol x_N)])/N$ for each combination of magnitude of $\\boldsymbol \\theta$ and model size. As the magnitude of $\\boldsymbol \\theta$ grows, so does the value of this metric.", fig.height=4}
p.max_q
```

There are $H + V + H*V$ terms in the sum that comprises $Q(\boldsymbol x)$, which leads to a natural split of the quantity into main effects, those terms based solely on $h_i$ and $v_j$ and interaction effects, those terms based on the product of $h_i v_j$. There are $H+V$ main effects terms and $H*V$ interaction effects terms. For $H$ and $V$ large enough, the number of interaction terms is growing much faster than $H + V$, meaning no matter the $C$, we can restrict the size of the left side of equation \ref{eq:stable} effectively by restricting the size of the interaction terms. This points to the interaction terms as potentially the more dangerous parameters for causing instability in the RBM simply because there are so many more of them.

```{r interaction size, include=FALSE}
expand.grid(H = 1:10, V = 1:10) %>%
  mutate(interaction_terms = H*V) %>%
  ggplot() +
  geom_contour(aes(H, V, z = interaction_terms/(H + V), colour = ..level..), bins = 20) +
  geom_contour(aes(H, V, z = interaction_terms/(H + V)), colour = "black", breaks = 1) +
  scale_colour_gradient("H*V/(H + V)", low = "yellow", high = "red")

### 2016-06-22 writing/draft.Rmd
beef up instability

###Instability
@schweinberger2011instability introduces a related concept, *instability*, which occurs when small changes in the natural parameters result in large changes in probability masses. This can be thought of as excessive sensitivity in the model. @schweinberger2011instability provides the following tool for detecting instability in the a discrete exponential family model.

\begin{defn}
A discrete exponential family distribution is \emph{unstable} if for any $C > 0$, however large, there exists $N_C > 0$ such that 
\begin{align}
\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] > CN \text{ for all } N > N_C \label{eq:unstable}
\end{align}

Conversely, a discrete exponential family distribution is \emph{stable} if there exists constants $C > 0$ and $N_C > 0$ such that 
\begin{align}
\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] \le CN \text{ for all } N > N_C. \label{eq:stable}
\end{align}
\end{defn}

Within the context of the RBM, $N$ is the number of nodes in the model, $H + V$. To ensure the model is stable, we must restrict $\boldsymbol \theta$ such that there exists constants $C > 0$ and $N_C > 0$ that force equation \ref{eq:stable} to hold.

\begin{align*}
\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] &= \max\limits_{\boldsymbol x \in \mathcal{X}}\left[ \sum\limits_{i = 1}^H\sum\limits_{j=1}^V \theta_{ij} h_i v_j + \sum\limits_{i = 1}^H\theta_{hi} h_i + \sum\limits_{j = 1}^V\theta_{vj} v_j \right] \\
&= \max\limits_{h_j, v_i \in \{0,1\}}\left[ \sum\limits_{i = 1}^H\sum\limits_{j=1}^V \theta_{ij} h_i v_j + \sum\limits_{i = 1}^H\theta_{hi} h_i + \sum\limits_{j = 1}^V\theta_{vj} v_j \right] \\
&\le C(H + V)
\end{align*}

There are $H + V + H*V$ terms in the sum that comprises $Q(\boldsymbol x)$, which leads to a natural split of the quantity into main effects, those terms based solely on $h_i$ and $v_j$ and interaction effects, those terms based on the product of $h_i v_j$. There are $H+V$ main effects terms and $H*V$ interaction effects terms. For $H$ and $V$ large enough, the number of interaction terms is growing much faster than $H + V$, meaning no matter the $C$, we can restrict the size of the left side of equation \ref{eq:stable} effectively by restricting the size of the interaction terms. This points to the interaction terms as potentially the more dangerous parameters for causing instability in the RBM simply because there are so many more of them.

```{r interaction size, include=FALSE}
expand.grid(H = 1:10, V = 1:10) %>%
  mutate(interaction_terms = H*V) %>%
  ggplot() +
  geom_contour(aes(H, V, z = interaction_terms/(H + V), colour = ..level..), bins = 20) +
  geom_contour(aes(H, V, z = interaction_terms/(H + V)), colour = "black", breaks = 1) +
  scale_colour_gradient("H*V/(H + V)", low = "yellow", high = "red")

@@

expand.grid(theta_h1 = seq(-5, 5, length.out = 5),
            theta_v1 = seq(-5, 5, length.out = 5),
            theta_11 = seq(-5, 5, length.out = 11)) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(gamma = sum(exp(possibles %*% t(data.matrix(.)))),
                mu_v1 = possibles[,"v1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_h1 = possibles[,"h1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_11 = possibles[,"theta11"] %*% exp(possibles %*% t(data.matrix(.))))) %>%
  mutate_each(funs(./gamma), starts_with("mu")) %>%
  select(-gamma) %>%
  gather(mu_parameter, value, mu_v1, mu_h1, mu_11) %>%
  ungroup() %>%
  mutate(theta_h1 = factor(theta_h1, labels = paste0("theta[h1]==", unique(theta_h1))), 
         theta_v1 = factor(theta_v1, labels = paste0("theta[v1]==", unique(theta_v1)))) %>%
  ggplot() +
  geom_line(aes(theta_11, value, colour = mu_parameter)) +
  facet_grid(theta_h1 ~ theta_v1, labeller = label_parsed) +
  xlab(expression(theta[11])) +
  scale_colour_discrete(expression(mu(theta)), labels = c(expression(theta[v1]), 
                                                          expression(theta[h1]), 
                                                          expression(theta[11])))

Secondly, we can look at $\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)]$ for this tiny model.

\begin{align*}
\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] &= \max\limits_{\boldsymbol x \in \mathcal{X}}\left[ \theta_{11} h_1 v_1 + \theta_{h_1}h_1 + \theta_{v_1} v_1 \right] \\
&=\max\left\{ \theta_{11} + \theta_{h_1} + \theta_{v_1}, \theta_{h_1}, \theta_{v_1}, 0  \right\} \\
\end{align*}

Figure \ref{fig:instab} shows $\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)]/N$ for various values of $\boldsymbol \theta$. Of course, we are interested in how this quantity grows as $N = H+V$ grows for various parameter values. We can see that this quantity is increasing for larger positive values of $\boldsymbol \theta$, which may indicate instability as $|\boldsymbol \theta|$ is large, agreeing with the indications about near-degeneracy from looking at $\mu(\boldsymbol \theta)$.

```{r instab, fig.cap="\\label{fig:instab}$(\\max\\limits_{\\boldsymbol x_N \\in \\mathcal{X}_N}[Q(\\boldsymbol x_N)] - \\min\\limits_{\\boldsymbol x_N \\in \\mathcal{X}_N}[Q(\\boldsymbol x_N)])/N$ for various values of $\\boldsymbol \\theta$ for the tiny example model. This quantity is increasing for larger positive values of $\\boldsymbol \\theta$.", fig.height = 4, fig.align='center'}
expand.grid(theta_h1 = seq(-5, 5, length.out = 5),
            theta_v1 = seq(-5, 5, length.out = 5),
            theta_11 = seq(-5, 5, length.out = 31)) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(q_max = max(possibles %*% t(data.matrix(.))) - min(possibles %*% t(data.matrix(.))))) %>%
  ungroup() %>%
  mutate(theta_h1 = factor(theta_h1, labels = paste0("theta[h1]==", unique(theta_h1))), 
         theta_v1 = factor(theta_v1, labels = paste0("theta[v1]==", unique(theta_v1)))) %>%
  ggplot() +
  geom_line(aes(theta_11, q_max/2)) +
  facet_grid(theta_h1 ~ theta_v1, labeller = label_parsed) +
  xlab(expression(theta[11])) +
  ylab(expression(frac(max(Q) - min(Q), H + V)))
```

@@

```{r uninterp, fig.cap="\\label{fig:uninterp}The difference between model expectations, E$\\left[\\boldsymbol X | \\boldsymbol \\theta\\right]$, and expectations given independence, E$\\left[\\boldsymbol X | \\emptyset\\right ]$ for a RBM with one visible and one hidden node.", fig.height = 4, fig.align='center', message=FALSE, cache=TRUE}
possibles <- stats(1, 1, type = "binary")

expand.grid(theta_h1 = seq(-5, 5, length.out = 5),
            theta_v1 = seq(-5, 5, length.out = 5),
            theta_11 = seq(-5, 5, length.out = 11)) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(gamma = sum(exp(possibles %*% t(data.matrix(.)))),
                mu_v1_dep = possibles[,"v1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_h1_dep = possibles[,"h1"] %*% exp(possibles %*% t(data.matrix(.))))) %>%
  mutate_each(funs(./gamma), starts_with("mu")) %>%
  select(-gamma) -> exp_dep

expand.grid(theta_h1 = seq(-5, 5, length.out = 5),
            theta_v1 = seq(-5, 5, length.out = 5),
            theta_11 = 0) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(gamma = sum(exp(possibles %*% t(data.matrix(.)))),
                mu_v1_indep = possibles[,"v1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_h1_indep = possibles[,"h1"] %*% exp(possibles %*% t(data.matrix(.))))) %>%
  mutate_each(funs(./gamma), starts_with("mu")) %>% ungroup() %>%
  select(-gamma, -theta_11) -> exp_indep

exp_dep %>%
  left_join(exp_indep) %>%
  mutate(mu_v1 = abs(mu_v1_dep - mu_v1_indep),
         mu_h1 = abs(mu_h1_dep - mu_h1_indep)) %>%
  gather(mu_parameter, value, mu_v1, mu_h1) %>%
  ungroup() %>%
  mutate(theta_h1 = factor(theta_h1, labels = paste0("theta[h1]==", unique(theta_h1))), 
         theta_v1 = factor(theta_v1, labels = paste0("theta[v1]==", unique(theta_v1)))) %>%
  ggplot() +
  geom_line(aes(theta_11, value, colour = mu_parameter)) +
  facet_grid(theta_h1 ~ theta_v1, labeller = label_parsed) +
  xlab(expression(theta[11])) + ylab(expression(group("|", E(group("", bold(X), "|")*bold(theta)) - E(group("", bold(X), "|")*plain(independence)), "|"))) +
  scale_colour_discrete(expression(bold(X)), labels = c(expression(v[1]), 
                                                          expression(h[1])))

Figure \ref{fig:uninterp} shows that the absolute difference between model expectations, E$\left[\boldsymbol X | \boldsymbol \theta\right]$, and expectations given independence, E$\left[\boldsymbol X | \emptyset\right ]$ for the RBM with one hidden and one visible node grows as the values of $\boldsymbol \theta$ aare farther from zero. This is a third indication that parameters of large magnetude could lead to model impropriety in an RBM using a model with only one hidden and one visible node.

@@

```{r instab_plots, fig.cap="\\label{fig:instab_plots}A numerical experiment looking at $(\\max\\limits_{\\boldsymbol x_N \\in \\mathcal{X}_N}[Q(\\boldsymbol x_N)] - \\min\\limits_{\\boldsymbol x_N \\in \\mathcal{X}_N}[Q(\\boldsymbol x_N)])/N$ for each combination of magnitude of $\\boldsymbol \\theta$ and model size. As the magnitude of $\\boldsymbol \\theta$ grows, so does the value of this metric.", fig.height=4}
p.max_q
```

@@

**TODO: Add proof that likelihood goes to infinity.**  
**TODO: Add marginalized Jing prior?**  
**TODO: Add comparison to contrastive divergence**

@@

>>>>

###Instability
@schweinberger2011instability introduces a related concept, *instability*, which occurs when small changes in the natural parameters result in large changes in probability masses. This can be characterized of as excessive sensitivity in the model, where a small change in $\boldsymbol x$ leads to a large change in the probability mass function. @schweinberger2011instability provides the following tool for detecting instability in the a discrete exponential family model.

\begin{defn}
A discrete exponential family distribution is \emph{unstable} if for any $C > 0$, however large, there exists $N_C > 0$ such that 
\begin{align}
\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] - \min\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] > CN \text{ for all } N > N_C \label{eq:unstable}
\end{align}

Conversely, a discrete exponential family distribution is \emph{stable} if there exists constants $C > 0$ and $N_C > 0$ such that 
\begin{align}
\max\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] - \min\limits_{\boldsymbol x_N \in \mathcal{X}_N}[Q(\boldsymbol x_N)] \le CN \text{ for all } N > N_C. \label{eq:stable}
\end{align}
\end{defn}

Within the context of the RBM, this definition concerns the joint distribution of both hidden and visible nodes, where $N$ is the number of nodes in the model, $H + V$. However, instability in terms of only the visible nodes is a more useful measure of model impropriety to be concerned with. To see this, consider the example of classifying images using an RBM. One would be interested in the maximum change in the likelihood when a single pixel is changed in the visibles as a proxy for instability in the model. Thus, we can instead concern ourselves with the notion of instability that deals with the maximum likelihood ratio of two images that differ in only one pixel.

\begin{align*}
&\max\left\{\log\frac{P_{\boldsymbol \theta}(v_1, \dots, v_V)}{P_{\boldsymbol \theta}(v^*_1, \dots, v^*_V)}: (v_1, \dots, v_V) \text{ and } (v^*_1, \dots, v^*_V) \in \{-1, 1\}^V \text{ differ by one component} \right\} \\
&\quad = \max\limits_{k \in 1, \dots, V}\left\{\log\frac{\sum\limits_{\boldsymbol h \in \{-1,1\}^H} \exp\left(
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right)}{\sum\limits_{\boldsymbol h \in \{-1,1\}^H} \exp\left(
\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right)}: v_k \not= v^*_k \right\} \\
& \quad \ge \max\limits_{k \in 1, \dots, V}\left\{\log\frac{\max\limits_{\boldsymbol h \in \{-1,1\}^H}\exp\left(
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right)}{2^H\max\limits_{\boldsymbol h \in \{-1,1\}^H}\exp\left(
\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right)}: v_k \not= v^*_k \right\} \\
&\quad = \max\limits_{k \in 1, \dots, V}\left\{\max\limits_{\boldsymbol h \in \{-1,1\}^H}\left\{
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right\}\right. \\
& \quad \quad \left. - \max\limits_{\boldsymbol h \in \{-1,1\}^H}
\left\{\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right\} - H\log2: v_k \not= v^*_k \right\} \\
&\quad \ge \max\limits_{\boldsymbol v \in \{-1,1\}^V}\max\limits_{\boldsymbol h \in \{-1,1\}^H}Q(\boldsymbol x) - \min\limits_{\boldsymbol v \in \{-1,1\}^V}\max\limits_{\boldsymbol h \in \{-1,1\}^H}Q(\boldsymbol x) - H\log 2
\end{align*}

Let $R(\boldsymbol \theta) \equiv \max_{\boldsymbol v} \max_{\boldsymbol h}Q(\boldsymbol x) - \min_{\boldsymbol v}\max_{\boldsymbol h}Q(\boldsymbol x) - H\log 2$. Then, if $R(\boldsymbol \theta) > C*V$, this implies that the maximum log-likelihood ratio of two images that differ in only one pixel is greater than $C$. So, by examining $R(\boldsymbol\theta)/V$, we can determine a lower bound for the maximum log-likelihood ratio of two images that differ in only one pixel, which if large is consistent with the definition of instability detailed above.

@@

expand.grid(theta_h1 = seq(-5, 5, length.out = 21),
            theta_v1 = seq(-5, 5, length.out = 21),
            theta_11 = seq(-5, 5, length.out = 5)) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(gamma = sum(exp(possibles %*% t(data.matrix(.)))),
                mu_v1 = possibles[,"v1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_h1 = possibles[,"h1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_11 = possibles[,"theta11"] %*% exp(possibles %*% t(data.matrix(.))))) %>%
  mutate_each(funs(./gamma), starts_with("mu")) %>%
  select(-gamma) %>%
  gather(mu_parameter, value, mu_v1, mu_h1, mu_11) %>%
  ungroup() %>%
  mutate(theta_11 = factor(theta_11, labels = paste0("theta[11]==", unique(theta_11))),
         mu_parameter = factor(mu_parameter, labels = paste0(gsub("_", "(theta[", unique(mu_parameter)), "])"))) %>%
  ggplot() +
  geom_tile(aes(theta_v1, theta_h1, fill = value))  +
  geom_contour(aes(theta_v1, theta_h1, z = value), colour = "black") +
  facet_grid(mu_parameter ~ theta_11, labeller = label_parsed) +
  xlab(expression(theta[v1])) +
  ylab(expression(theta[h1])) +
  scale_fill_gradient2("Mean-parameterization", low = "#af8dc3", high = "#7fbf7b", midpoint = .5, limits = c(0, 1)) +
  theme(aspect.ratio = 1)
```

Secondly, we can look at $R(\boldsymbol \theta)$ for this tiny model, which gives us a lower bound for the loglikelihood ratio of the smallest possible change in $\boldsymbol x$, indicating instability.

\begin{align*}
R(\boldsymbol \theta) &= \max_{\boldsymbol v} \max_{\boldsymbol h}Q(\boldsymbol x) - \min_{\boldsymbol v}\max_{\boldsymbol h}Q(\boldsymbol x) - H\log 2\\
&= \max_{\boldsymbol v} \max_{\boldsymbol h}\left[ \theta_{11} h_1 v_1 + \theta_{h_1}h_1 + \theta_{v_1} v_1 \right] -  \min_{\boldsymbol v}\max_{\boldsymbol h}\left[ \theta_{11} h_1 v_1 + \theta_{h_1}h_1 + \theta_{v_1} v_1 \right] -\log 2\\
&=\max\left\{ \theta_{11} + \theta_{h_1} + \theta_{v_1}, \theta_{h_1}, \theta_{v_1}, 0  \right\} \\
&\qquad - \min\left\{\theta_{v_1} + \max\left\{\theta_{11} + \theta_{h1}, -\theta_{11} - \theta_{h1}\right\}, -\theta_{v_1} + \max\left\{-\theta_{11} + \theta_{h1}, \theta_{11} - \theta_{h1}\right\}\right\} - \log2\\
\end{align*}

Figure \ref{fig:instab} shows $R(\boldsymbol \theta)/V$ for various values of $\boldsymbol \theta$. We can see that this quantity is increasing for larger magnitudes of $\boldsymbol \theta$, especially for large values of $\theta_11$. This may indicate instability as $|\boldsymbol \theta|$ is large, agreeing with the indications about near-degeneracy from looking at $\mu(\boldsymbol \theta)$.

```{r instab, fig.cap="\\label{fig:instab}$R(\\boldsymbol \\theta)/V$ for various values of $\\boldsymbol \\theta$ for the tiny example model. This quantity is increasing for larger positive values of $\\boldsymbol \\theta$.", fig.height = 4, fig.align='center'}
expand.grid(theta_h1 = seq(-5, 5, length.out = 21),
            theta_v1 = seq(-5, 5, length.out = 21),
            theta_11 = seq(-5, 5, length.out = 5)) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(Q = cbind(possibles, possibles %*% t(data.matrix(.))))) %>%
  group_by(theta_v1, theta_h1, theta_11, Q.v1) %>%
  do(data.frame(maxq = max(.$Q.V4))) %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(q_max = max(.$maxq) - min(.$maxq) - log(2))) %>%
  ungroup() %>%
  mutate(theta_11 = factor(theta_11, labels = paste0("theta[11]==", unique(theta_11)))) %>%
  ggplot() +
  geom_tile(aes(theta_v1, theta_h1, fill = q_max)) +
  geom_contour(aes(theta_v1, theta_h1, z = q_max), colour = "black") +
  facet_grid(. ~ theta_11, labeller = label_parsed) +
  xlab(expression(theta[v1])) +
  ylab(expression(theta[h1])) +
  scale_fill_gradient(expression(frac(R(theta), V)), low = "yellow", high = "red") +
  theme(aspect.ratio = 1)
```

@@

```{r uninterp, fig.cap="\\label{fig:uninterp}The difference between model expectations, E$\\left[\\boldsymbol X | \\boldsymbol \\theta\\right]$, and expectations given independence, E$\\left[\\boldsymbol X | \\emptyset\\right ]$ for a RBM with one visible and one hidden node.", fig.height = 4, fig.align='center', message=FALSE, cache=TRUE, warning=FALSE}
expand.grid(theta_h1 = seq(-5, 5, length.out = 21),
            theta_v1 = seq(-5, 5, length.out = 21),
            theta_11 = seq(-5, 5, length.out = 5)) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(gamma = sum(exp(possibles %*% t(data.matrix(.)))),
                mu_v1_dep = possibles[,"v1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_h1_dep = possibles[,"h1"] %*% exp(possibles %*% t(data.matrix(.))))) %>%
  mutate_each(funs(./gamma), starts_with("mu")) %>%
  select(-gamma) -> exp_dep

expand.grid(theta_h1 = seq(-5, 5, length.out = 21),
            theta_v1 = seq(-5, 5, length.out = 21),
            theta_11 = 0) %>%
  data.frame() %>%
  group_by(theta_v1, theta_h1, theta_11) %>%
  do(data.frame(gamma = sum(exp(possibles %*% t(data.matrix(.)))),
                mu_v1_indep = possibles[,"v1"] %*% exp(possibles %*% t(data.matrix(.))),
                mu_h1_indep = possibles[,"h1"] %*% exp(possibles %*% t(data.matrix(.))))) %>%
  mutate_each(funs(./gamma), starts_with("mu")) %>% ungroup() %>%
  select(-gamma, -theta_11) -> exp_indep

exp_dep %>%
  left_join(exp_indep) %>%
  mutate(mu_v1 = abs(mu_v1_dep - mu_v1_indep),
         mu_h1 = abs(mu_h1_dep - mu_h1_indep)) %>%
  gather(mu_parameter, value, mu_v1, mu_h1) %>%
  ungroup() %>%
  mutate(theta_11 = factor(theta_11, labels = paste0("theta[11]==", unique(theta_11))),
         mu_parameter = factor(mu_parameter, labels = paste0(gsub("_", "(theta[", unique(mu_parameter)), "])"))) %>%
  ggplot() +
  geom_tile(aes(theta_v1, theta_h1, fill = value))  +
  geom_contour(aes(theta_v1, theta_h1, z = value), colour = "black") +
  facet_grid(mu_parameter ~ theta_11, labeller = label_parsed) +
  xlab(expression(theta[v1])) +
  ylab(expression(theta[h1])) +
  scale_fill_gradient(expression(group("|", E(group("", bold(X), "|")*bold(theta)) - E(group("", bold(X), "|")*plain(independence)), "|")), low = "yellow", high = "red", limits = c(0, 1)) +
  theme(aspect.ratio = 1)
```

Figure \ref{fig:uninterp} shows that the absolute difference between model expectations, E$\left[\boldsymbol X | \boldsymbol \theta\right]$, and expectations given independence, E$\left[\boldsymbol X | \emptyset\right ]$ for the RBM with one hidden and one visible node grows as the values of $\boldsymbol \theta$ are farther from zero, especially for large magnitudes of $\theta_11$. This is a third indication that parameters of large magnitude could lead to model impropriety in an RBM using a model with only one hidden and one visible node.

@@

@@

**TODO: Add proof that likelihood goes to infinity.**  
**TODO: Add comparison to contrastive divergence**

@@

##Center of the universe

##Incomplete cases
